---
title: "R Notebook"
output: html_notebook
---

### Problem 1

```{r}
equity_process1 <- function(mu, sigma, r, A0, B0, t, m){
  dt <- t/m
  time <- seq(0, t, length.out=(m + 1) )
  Z <- rnorm(m)
  Wt <- c(0, cumsum(sqrt(dt) * Z))
  
  A <- A0 * exp((mu - (sigma^2)/2) * time + sigma * Wt)
  B <- B0 * exp(r * time)
  return(pmax(A - B,0))
}

```

#(1a) Sample paths
```{r}
E <- equity_process1(mu=0.1, sigma=0.2, r=0.1, A0=1000, B0=900, t=10 , m= 1e2)
time <- seq(0, t =10, length.out = length(E))
plot(time, E, type="l")
```

#(1b) Default probability
```{r}
sim <- replicate(1e4, equity_process1(mu=0.1, sigma=0.2, r=0.1, A0=1000, B0=900, t=5 , m= 1e2))
mean(sim[nrow(sim),] == 0)

```

#(1c) Adding jump processes


```{r}
equity_process3 <- function(mu,t,sigma, n, A0, B0 ,r, nsim, lambda){
  dt <- t/n
  Z <- matrix(rnorm(n*nsim), ncol = nsim)
  
  A <- matrix(0,nrow = n+1, ncol = nsim)
  B <- matrix(0,nrow = n+1, ncol = nsim)
  E <- matrix(0,nrow = n+1, ncol = nsim)
  jump <- function(dNt){
    if(dNt >= 1){ return(sum(runif(dNt,-500,500)))} 
    else(return(0))
  }
  
  A[1, ]<- A0
  B[1,] <- B0
  for(i in 1:n){
    dNt <- rpois(nsim, lambda*dt)
    dJt <- vapply(dNt, jump, numeric(1))
    A[i+1,] <- A[i,] + A[i,] *(mu*dt + sigma*Z[i,]*sqrt(dt)) + dJt
    B[i+1,] <- B[i,]+ r*B[i,]*dt
    
    E[i+1,] <- A[i+1,] - B[i+1,]
    if( E[i+1,] <= rep(0,n)){
      E[(i+2):(m+1),] <- 0
      break
      
    }
  }

  return(E)
  
}
```

```{r}
E3 <- equity_process3(mu = 0.1, t = 5, sigma = 0.2, n =1e3, A0 = 1000, B0 = 900, r =0.1, nsim = 1e3, lambda = 1.2)
mean(E3[nrow(E3),] <0)
```

### Problem 2
#(2a) The Heston Model.
```{r}
S0 <- 50; r <- 0.03; a <- 1; b <- 0.2; sigma <- 0.6; theta0 <- 0.3; T <- 1; K <- 50; rho <- 0.6; m <- 50

heston_path <- function(S0, r, a, b, sigma, theta0, T, K, rho, m){
  dt <- T/m
  
  S <- numeric(m+1)
  theta <- numeric(m+1)
  S[1] <- S0
  theta[1] <- theta0
  
  Z1 <- rnorm(m)
  Z2 <- rnorm(m)
  #giải cholesky
  
  Z_1 <- Z1
  Z_2 <- rho*Z1 + sqrt(1-rho^2)*Z2
  
  for (i in 1:m) {
    theta[i] <- max(theta[i],0)
    theta[i+1] <- theta[i] + a*(b-theta[i])*dt + sigma*sqrt(theta[i])*sqrt(dt)*Z_1[i]
    S[i+1] <- S[i] + r*S[i]*dt + sqrt(theta[i])*S[i]*sqrt(dt)*Z_2[i]
  }
  euro <- max(S[m+1] - K, 0)
  avg <- max(mean(S[2:(m+1)]) - K, 0)
  price_euro <- euro*exp(-r*T)
  price_avg <- avg*exp(-r*T)
  
  return(c(price_euro, price_avg))
}

nsim <- 1e5
sim <- replicate(nsim, heston_path(S0, r, a, b, sigma, theta0, T, K, rho, m))

euro_mean <- mean(sim[1,])
euro_se   <- sd(sim[1,]) / sqrt(nsim)
aver_mean <- mean(sim[2,])
aver_se   <- sd(sim[2,]) / sqrt(nsim)

cat("European call option price:", euro_mean, "| SE:", euro_se, "\n",
    "Average call option price:", aver_mean, "| SE:", aver_se, "\n")
```

#(2b) Variance Reduction: Antithetic

```{r}
S0 <- 50; r <- 0.03; a <- 1; b <- 0.2; sigma <- 0.6; theta0 <- 0.3; T <- 1; K <- 50; rho <- 0.6; m <- 50

heston_path_anti <- function(S0, r, a, b, sigma, theta0, T, K, rho, m){
  dt <- T/m
  
  S <- numeric(m+1)
  S_anti <- numeric(m+1)
  theta <- numeric(m+1)
  theta_anti <- numeric(m+1)
  
  S[1] <- S0
  S_anti[1] <- S0
  theta[1] <- theta0
  theta_anti[1] <- theta0
  
  Z1 <- rnorm(m)
  Z2 <- rnorm(m)
  # tính cholesky
  Z_1 <- Z1
  Z_1_anti <- -Z_1
  
  Z_2 <- rho*Z1 + sqrt(1-rho^2)*Z2
  Z_2_anti <- -Z_2
  
  for (i in 1:m) {
    theta[i] <- max(theta[i],0)
    
    theta[i+1] <- theta[i] + a*(b-theta[i])*dt + sigma*sqrt(theta[i])*sqrt(dt)*Z_1[i]
    S[i+1] <- S[i] + r*S[i]*dt + sqrt(theta[i])*S[i]*sqrt(dt)*Z_2[i]
    
    #Anti path
    theta_anti[i] <- max(theta_anti[i],0)
    
    theta_anti[i+1] <- theta_anti[i] + a*(b-theta_anti[i])*dt + sigma*sqrt(theta_anti[i])*sqrt(dt)*Z_1_anti[i]
    S_anti[i+1] <- S_anti[i] + r*S_anti[i]*dt + sqrt(theta_anti[i])*S_anti[i]*sqrt(dt)*Z_2_anti[i]
  }

  price_euro <- max(S[m+1] - K, 0)*exp(-r*T)
  price_avg <- max(mean(S[2:(m+1)]) - K, 0)*exp(-r*T)
  price_euro_anti <- max(S_anti[m+1] - K, 0)*exp(-r*T)
  price_avg_anti <- max(mean(S_anti[2:(m+1)]) - K, 0)*exp(-r*T)
  
  euro_VR <- (price_euro + price_euro_anti)/2
  avg_VR <- (price_avg + price_avg_anti)/2
  
  return(c(euro_VR, avg_VR))
}

nsim <- 1e5
simlist_VR <- replicate(nsim, heston_path_anti(S0, r, a, b, sigma, theta0, T, K, rho, m))
price_euro_VR <- mean(simlist_VR[1, ])
price_avg_VR <- mean(simlist_VR[2, ])

se_euro_VR <- sd(simlist_VR[1, ])/sqrt(nsim)
se_avg_VR <- sd(simlist_VR[2, ])/sqrt(nsim)

cat("European call option price:", price_euro_VR, "| SE:", se_euro_VR, "\n",
    "Average call option price:", price_avg_VR, "| SE:", se_avg_VR, "\n")
```
#report for comparison:
Report:

While the Euler scheme introduces discretization error (which is controlled by the number of steps), the Antithetic Variates method focuses on reducing the statistical sampling error (captured by the SE).

Path Dependency: The significant reduction in SE (~36%) for the Asian Call highlights the effectiveness of antithetic paths in smoothing out the average price volatility throughout the simulation period.

Computational Efficiency: By utilizing antithetic pairs, we achieved a substantially higher level of precision (tighter confidence intervals) without the need to increase the number of paths, thereby optimizing the computation.


#(2c) Delta hedging by Euler scheme.
```{r}
Eu_Delta_Heston <- function(nsim, S0, r, a, b, sigma, theta0, T, K, rho, m) {
  dt <- T/m
  L <- t(chol(matrix(c(1, rho, rho, 1), 2, byrow = TRUE)))
  
  # Khởi tạo ma trận
  S <- matrix(0, m + 1, nsim)
  theta <- matrix(0, m + 1, nsim)
  Delta_path <- matrix(0, m + 1, nsim) 
  
  S[1, ] <- S0
  theta[1, ] <- theta0
  Delta_path[1, ] <- 1 # Hint: Delta_0 = 1
  
  for (t in 1:m) {
    Z_raw <- matrix(rnorm(2 * nsim), nrow = 2)
    Z <- L %*% Z_raw 
    
    theta[t, ] <- pmax(theta[t, ], 0)
    
    theta[t+1, ] <- theta[t, ] + a*(b - theta[t, ])*dt + sigma*sqrt(theta[t, ])*Z[1, ]*sqrt(dt)
    
    S[t+1, ] <- S[t, ] + r*S[t, ]*dt + sqrt(theta[t, ])*S[t, ]*Z[2, ]*sqrt(dt)
    
    Delta_path[t+1, ] <- Delta_path[t, ] * (1 + r*dt + sqrt(theta[t, ])*Z[2, ]*sqrt(dt))
  }
  
  # Lấy giá trị tại bước cuối T
  ST <- S[m+1, ]
  Delta_T <- Delta_path[m+1, ]
  
  # Indicator(ST >= K) trong R trả về 1 nếu đúng, 0 nếu sai
  delta_estimates <- exp(-r * T) * (ST >= K) * Delta_T
  
  res_delta <- mean(delta_estimates)
  res_se <- sd(delta_estimates) / sqrt(nsim)
  
  return(data.frame(Delta = res_delta, SE = res_se))
}

result_delta <- Eu_Delta_Heston(nsim=1e4, S0=50, r=0.03, a=1, b=0.2, sigma=0.6, theta0=0.3, T=1, K=50, rho=0.6, m=50)
print(result_delta)
```

